---
title: "ECON 1001 Spring 2026: Course Grades Report"
subtitle: "INTERNAL — NOT FOR DISTRIBUTION"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    number_sections: false
    fig_caption: false
header-includes:
  - \usepackage{mathpazo}
fontsize: 11pt
geometry: "margin=1.1in"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(kableExtra)
library(knitr)

score_to_gp <- function(score) {
  case_when(
    is.na(score)   ~ NA_real_,
    score > 92.2   ~ 4.00,
    score > 89.3   ~ 3.67,
    score > 86.3   ~ 3.33,
    score > 82.3   ~ 3.00,
    score > 79.3   ~ 2.67,
    score > 76.3   ~ 2.33,
    score > 72.3   ~ 2.00,
    score > 69.3   ~ 1.67,
    score > 66.3   ~ 1.33,
    score > 62.3   ~ 1.00,
    TRUE           ~ 0.00
  )
}

gp_to_letter <- function(gp) {
  case_when(
    gp >= 4.00 ~ "A",
    gp >= 3.67 ~ "A-",
    gp >= 3.33 ~ "B+",
    gp >= 3.00 ~ "B",
    gp >= 2.67 ~ "B-",
    gp >= 2.33 ~ "C+",
    gp >= 2.00 ~ "C",
    gp >= 1.67 ~ "C-",
    gp >= 1.33 ~ "D+",
    gp >= 1.00 ~ "D",
    gp >= 0.00 ~ "F",
    TRUE       ~ NA_character_
  )
}

make_stats <- function(x) {
  tibble(
    Statistic = c("N", "Mean", "Std dev", "Minimum",
                  "10th pctile", "25th pctile", "Median",
                  "75th pctile", "90th pctile", "Maximum"),
    Value = c(
      as.character(sum(!is.na(x))),
      formatC(mean(x,              na.rm = TRUE), format = "f", digits = 1),
      formatC(sd(x,                na.rm = TRUE), format = "f", digits = 1),
      formatC(min(x,               na.rm = TRUE), format = "f", digits = 1),
      formatC(quantile(x, .10,     na.rm = TRUE), format = "f", digits = 1),
      formatC(quantile(x, .25,     na.rm = TRUE), format = "f", digits = 1),
      formatC(median(x,            na.rm = TRUE), format = "f", digits = 1),
      formatC(quantile(x, .75,     na.rm = TRUE), format = "f", digits = 1),
      formatC(quantile(x, .90,     na.rm = TRUE), format = "f", digits = 1),
      formatC(max(x,               na.rm = TRUE), format = "f", digits = 1)
    )
  )
}

# Students who received a +2 podcast bonus for excellent podcast (10 → 12)
EXCELLENT_PODCAST_SIDS <- c(
  "js5604", "jku3",   "eas426", "fkb8",   "gg850",
  "ls1772", "ak2594", "zz658",  "ma2570", "jl3390",
  "mcm539", "cf981",  "mq148",  "ps1436", "aks212"
)

df <- readRDS(here("src/grading/course/build/course_scores.rds")) %>%
  mutate(
    used_alt    = course_score_s2 > course_score_s1,
    bumped      = !is.na(adjusted_gp) & !is.na(gp) & adjusted_gp > gp,
    letter      = gp_to_letter(adjusted_gp),
    letter      = factor(letter,
                         levels = c("A","A-","B+","B","B-","C+","C","C-","D+","D","F"))
  )
```

### Section 1: Overview

```{r overview}
n_total      <- nrow(df)
n_graded     <- sum(!is.na(df$final_exam))
n_missing    <- sum(is.na(df$final_exam))
n_no_podcast <- sum(df$podcast_unsubmitted, na.rm = TRUE)
n_bumped     <- sum(df$bumped,   na.rm = TRUE)
n_failing    <- sum(df$letter == "F", na.rm = TRUE)
```

`r n_total` students enrolled. `r n_graded` have a final exam score; `r n_missing` do not (absent or pending). `r n_no_podcast` had no podcast submission (score set to 0; see Section 7). `r n_bumped` received a bump (final exam GP > course GP and final $\geq$ midterm average + 10). `r n_failing` received an F.

---

### Section 2: Component Score Summary

```{r component-summary}
has_final <- df %>% filter(!is.na(final_exam))

component_summary <- has_final %>%
  select(
    `PS avg (out of 10)`    = psavg,
    `Podcast (out of 10)`   = podcast,
    `Midterm 1`             = midterm1,
    `Midterm 2`             = midterm2,
    `Final exam (curved)`   = final_exam,
    `Course score`          = course_score
  ) %>%
  pivot_longer(everything(), names_to = "Component", values_to = "Score") %>%
  group_by(Component) %>%
  summarise(
    N      = sum(!is.na(Score)),
    Mean   = round(mean(Score,   na.rm = TRUE), 1),
    Median = round(median(Score, na.rm = TRUE), 1),
    SD     = round(sd(Score,     na.rm = TRUE), 1),
    Min    = round(min(Score,    na.rm = TRUE), 1),
    Max    = round(max(Score,    na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  arrange(match(Component, c("PS avg (out of 10)", "Podcast (out of 10)",
                              "Midterm 1", "Midterm 2",
                              "Final exam (curved)", "Course score")))

kable(component_summary, booktabs = TRUE, linesep = "",
      caption = "Summary of Score Components") %>%
  kable_styling(latex_options = "hold_position", full_width = FALSE, position = "center")
```

---

### Section 3: Course Score Distribution

```{r course-score-stats}
kable(make_stats(has_final$course_score),
      booktabs = TRUE, linesep = "", align = "lr",
      col.names = c("Statistic", "Value"),
      caption = "Course Score Distribution (students with final exam)") %>%
  kable_styling(latex_options = "hold_position", full_width = FALSE, position = "center")
```

```{r course-score-hist, fig.height=3.2, out.width="88%", fig.align="center"}
ggplot(has_final, aes(x = course_score)) +
  geom_histogram(binwidth = 2, fill = "#2c7fb8", color = "white", boundary = 0) +
  geom_vline(xintercept = median(has_final$course_score, na.rm = TRUE),
             linetype = "dashed", color = "black") +
  labs(x = "Course score", y = "Count",
       caption = paste0("Dashed line = median (",
                        round(median(has_final$course_score, na.rm = TRUE), 1), ")")) +
  theme_minimal(base_size = 11)
```

---

### Section 4: Letter Grade Distribution

```{r grade-dist}
grade_dist <- df %>%
  filter(!is.na(letter)) %>%
  count(letter, .drop = FALSE) %>%
  mutate(Pct = round(n / sum(n) * 100, 1))

kable(grade_dist, booktabs = TRUE, linesep = "",
      col.names = c("Grade", "Count", "%"),
      caption = "Letter Grade Distribution") %>%
  kable_styling(latex_options = "hold_position", full_width = FALSE, position = "center")
```

---

### Section 5: Bump Rule Recipients

If a student's final exam GP is higher than their course GP **and** their final exam score is at least 10 points above their midterm average, they receive the final exam GP as their course grade instead.

```{r bumped}
bumped_tbl <- df %>%
  filter(bumped) %>%
  transmute(
    Student        = student,
    `Midterm avg`  = round(midterm_avg, 1),
    `Final exam`   = round(final_exam,  1),
    `Course score` = round(course_score, 1),
    `Pre-bump GP`  = gp,
    `Final GP`     = adjusted_gp,
    Grade          = letter
  ) %>%
  arrange(desc(`Final exam`))

if (nrow(bumped_tbl) == 0) {
  cat("No students received a bump.\n")
} else {
  kable(bumped_tbl, booktabs = TRUE, linesep = "",
        caption = paste0("Students Who Received a Bump (n = ", nrow(bumped_tbl), ")")) %>%
    kable_styling(latex_options = c("hold_position", "scale_down"),
                  full_width = FALSE, position = "center")
}
```

---

### Section 6: Students with No Final Exam Score

```{r no-final}
no_final <- df %>%
  filter(is.na(final_exam)) %>%
  transmute(
    Student   = student,
    Midterm1  = round(midterm1, 1),
    Midterm2  = round(midterm2, 1),
    `PS avg`  = round(psavg, 1)
  )

if (nrow(no_final) == 0) {
  cat("All students have a final exam score.\n")
} else {
  kable(no_final, booktabs = TRUE, linesep = "",
        caption = paste0("Students with No Final Exam Score (n = ", nrow(no_final), ")")) %>%
    kable_styling(latex_options = "hold_position", full_width = FALSE, position = "center")
}
```

---

### Section 7: Unsubmitted Podcast — Student Detail

These students had no podcast submission and received a score of 0. Their course scores reflect the 0.

```{r no-podcast}
no_podcast <- df %>%
  filter(podcast_unsubmitted) %>%
  transmute(
    Student            = student,
    `PS avg (drop 2)`  = round(psavg,        1),
    `MT1 curved`       = round(midterm1,      1),
    `MT2 curved`       = round(midterm2,      1),
    `Final curved`     = round(final_exam,    1),
    Podcast            = podcast,
    `Course score`     = round(course_score,  1),
    Grade              = letter
  ) %>%
  arrange(Student)

if (nrow(no_podcast) == 0) {
  cat("All students have a podcast submission.\n")
} else {
  kable(no_podcast, booktabs = TRUE, linesep = "",
        caption = paste0("Unsubmitted Podcast — Score set to 0 (n = ", nrow(no_podcast), ")")) %>%
    kable_styling(latex_options = c("hold_position", "scale_down"),
                  full_width = FALSE, position = "center", font_size = 8)
}
```

---

### Section 8: Excellent Podcast Recipients — Grade Impact

Students who received a +2 podcast bonus (10 → 12) for excellent podcasts, as judged by graders. The bonus adds 1 point to the course score. The table flags whether this pushed anyone across a grade boundary.

```{r excellent-podcast}
# Counterfactual: podcast score was 10 instead of 12 → course score 1 pt lower
excellent <- df %>%
  filter(sis_user_id %in% EXCELLENT_PODCAST_SIDS) %>%
  mutate(
    course_score_counter = course_score - 1,
    gp_counter           = score_to_gp(course_score_counter),
    grade_changed        = !is.na(gp) & gp != gp_counter
  ) %>%
  transmute(
    Student              = student,
    `Course score`       = round(course_score, 1),
    `Score w/o bonus`    = round(course_score_counter, 1),
    Grade                = letter,
    `Grade w/o bonus`    = factor(gp_to_letter(gp_counter),
                                  levels = levels(letter)),
    `Grade changed?`     = if_else(grade_changed, "YES", "no")
  ) %>%
  arrange(desc(`Grade changed?`), Student)

kable(excellent, booktabs = TRUE, linesep = "",
      caption = "Excellent Podcast Recipients: Grade Impact of +2 Podcast Bonus") %>%
  kable_styling(latex_options = c("hold_position", "scale_down"),
                full_width = FALSE, position = "center")
```

---

### Section 9: Failing Students — Detail

```{r failing}
failing_tbl <- df %>%
  filter(letter == "F") %>%
  transmute(
    Student        = student,
    `PS avg`       = round(psavg,        1),
    `Midterm 1`    = round(midterm1,     1),
    `Midterm 2`    = round(midterm2,     1),
    `Final`        = round(final_exam,   1),
    `Course score` = round(course_score, 1)
  ) %>%
  arrange(`Course score`)

if (nrow(failing_tbl) == 0) {
  cat("No failing students.\n")
} else {
  kable(failing_tbl, booktabs = TRUE, linesep = "",
        caption = paste0("Failing Students (n = ", nrow(failing_tbl), ")")) %>%
    kable_styling(latex_options = c("hold_position", "scale_down"),
                  full_width = FALSE, position = "center")
}
```

---

### Section 10: Flags and Action Items

- **Missing final exam:** `r n_missing` student(s) have no final exam score and need follow-up.
- **Unsubmitted podcast:** `r n_no_podcast` student(s) had no podcast submission; scored 0 (see Section 7).
- **Failing:** `r n_failing` student(s) received an F.
- **Bump rule:** `r n_bumped` student(s) had their grade bumped by the final-exam-improvement rule.

---

### Section 11: Full Grade Roster

```{r full-roster}
roster <- df %>%
  arrange(desc(course_score)) %>%
  transmute(
    Student            = str_trunc(student, 35),
    `PS avg (drop 2)`  = round(psavg,        1),
    `MT1 curved`       = round(midterm1,      1),
    `MT2 curved`       = round(midterm2,      1),
    `Final curved`     = round(final_exam,    1),
    Podcast            = podcast,
    `Course score`     = round(course_score,  1),
    Grade              = letter
  )

kable(roster, booktabs = TRUE, linesep = "",
      caption = "Full Grade Roster, Ranked by Course Score") %>%
  kable_styling(latex_options = c("hold_position", "scale_down"),
                full_width = FALSE, position = "center",
                font_size = 8)
```
