---
title: "ECON 1001 — Final Exam: Internal Grading Note"
subtitle: "Spring 2026"
date: "`r format(Sys.Date(), '%B %d, %Y')`  |  INTERNAL — NOT FOR DISTRIBUTION"
output:
  pdf_document:
    number_sections: false
    fig_caption: false
header-includes:
  - \usepackage{mathpazo}
geometry: "margin=1.1in"
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(knitr)
library(kableExtra)
library(here)

# ── Configuration ──────────────────────────────────────────────
MC_MAX      <- 27
FR_MAX      <- 24
EXAM_MAX    <- 132   # 4 * MC_MAX + FR_MAX
PRELIMINARY <- FALSE

TARGET_MEDIAN <- 84.0
TARGET_MAX    <- 100.0
# ───────────────────────────────────────────────────────────────

mc  <- readRDS(here("src/grading/final-exam/build/final_exam_mc_clean.rds"))
fr  <- readRDS(here("src/grading/final-exam/build/final_exam_fr_clean.rds"))
fe  <- readRDS(here("src/grading/final-exam/build/final_exam_curved.rds"))

mc_graded <- mc %>% filter(mc_status == "Graded")
fr_graded <- fr %>% filter(fr_status == "Graded")

# Reconstruct curve parameters (mirrors 03_combine_and_curve.R)
raw_median <- median(fe$raw_score)
raw_max    <- max(fe$raw_score)
B_param    <- (TARGET_MAX - TARGET_MEDIAN) / (raw_max - raw_median)
A_param    <- TARGET_MEDIAN - B_param * raw_median

make_stats <- function(x, digits = 1, label_max = NULL) {
  rows <- tibble(
    Statistic = c("N", "Mean", "Std dev", "Minimum",
                  "10th percentile", "25th percentile",
                  "Median", "75th percentile",
                  "90th percentile", "Maximum"),
    Value = c(
      as.character(sum(!is.na(x))),
      formatC(mean(x,             na.rm = TRUE), format = "f", digits = digits),
      formatC(sd(x,               na.rm = TRUE), format = "f", digits = digits),
      formatC(min(x,              na.rm = TRUE), format = "f", digits = digits),
      formatC(quantile(x, .10,   na.rm = TRUE), format = "f", digits = digits),
      formatC(quantile(x, .25,   na.rm = TRUE), format = "f", digits = digits),
      formatC(median(x,           na.rm = TRUE), format = "f", digits = digits),
      formatC(quantile(x, .75,   na.rm = TRUE), format = "f", digits = digits),
      formatC(quantile(x, .90,   na.rm = TRUE), format = "f", digits = digits),
      formatC(max(x,              na.rm = TRUE), format = "f", digits = digits)
    )
  )
  if (!is.null(label_max))
    rows <- add_row(rows,
                    Statistic = paste0("Perfect scores (= ", label_max, ")"),
                    Value     = as.character(sum(x >= label_max, na.rm = TRUE)))
  rows
}

make_stats_dual <- function(pts_vec, max_val, digits = 1) {
  pct_vec <- pts_vec / max_val * 100
  fmtC <- function(x, d = digits) formatC(x, format = "f", digits = d)
  tibble(
    Statistic = c("N", "Mean", "Std dev", "Minimum",
                  "10th percentile", "25th percentile",
                  "Median", "75th percentile",
                  "90th percentile", "Maximum",
                  paste0("Perfect scores (= ", max_val, " pts)")),
    Points = c(
      as.character(sum(!is.na(pts_vec))),
      fmtC(mean(pts_vec,           na.rm = TRUE)),
      fmtC(sd(pts_vec,             na.rm = TRUE)),
      fmtC(min(pts_vec,            na.rm = TRUE)),
      fmtC(quantile(pts_vec, .10,  na.rm = TRUE)),
      fmtC(quantile(pts_vec, .25,  na.rm = TRUE)),
      fmtC(median(pts_vec,         na.rm = TRUE)),
      fmtC(quantile(pts_vec, .75,  na.rm = TRUE)),
      fmtC(quantile(pts_vec, .90,  na.rm = TRUE)),
      fmtC(max(pts_vec,            na.rm = TRUE)),
      as.character(sum(pts_vec >= max_val, na.rm = TRUE))
    ),
    `%` = c(
      "",
      fmtC(mean(pct_vec,           na.rm = TRUE)),
      fmtC(sd(pct_vec,             na.rm = TRUE)),
      fmtC(min(pct_vec,            na.rm = TRUE)),
      fmtC(quantile(pct_vec, .10,  na.rm = TRUE)),
      fmtC(quantile(pct_vec, .25,  na.rm = TRUE)),
      fmtC(median(pct_vec,         na.rm = TRUE)),
      fmtC(quantile(pct_vec, .75,  na.rm = TRUE)),
      fmtC(quantile(pct_vec, .90,  na.rm = TRUE)),
      fmtC(max(pct_vec,            na.rm = TRUE)),
      ""
    )
  )
}

```

`r if (PRELIMINARY) "\\textbf{PRELIMINARY:} Based on partial free-response data. Curve parameters will be recalibrated once all FR is graded.\n\n---\n"`

## Overview

The final exam consisted of **`r MC_MAX` multiple-choice questions** (4 pts each) and **6 free-response parts** (4 pts each), for a maximum of **`r EXAM_MAX` points**.

```{r overview-table}
n_enrolled <- nrow(mc)
n_mc       <- sum(mc$mc_status == "Graded")
n_fr       <- sum(fr$fr_status == "Graded")
n_both     <- nrow(fe)
n_mc_only  <- sum(mc$mc_status == "Graded" & !(mc$sis_user_id %in% fe$sis_user_id))

tibble(
  Item  = c("Enrolled", "MC graded", "FR graded", "Both graded (curved)",
            "FR ungraded", "FR missing"),
  Count = c(n_enrolled, n_mc, n_fr, n_both,
            sum(fr$fr_status == "Ungraded"),
            sum(fr$fr_status == "Missing"))
) %>%
  kable(booktabs = TRUE, linesep = "", align = "lc") %>%
  kable_styling(latex_options = "hold_position", full_width = FALSE, position = "center")
```

---

## MC Scores (out of `r MC_MAX`)

Based on **`r n_mc`** students with MC graded.

```{r mc-stats}
kable(make_stats(mc_graded$mc_score, label_max = MC_MAX, digits = 1),
      booktabs = TRUE, linesep = "", align = "lr",
      col.names = c("Statistic", "Value")) %>%
  kable_styling(latex_options = "hold_position", full_width = FALSE, position = "center")
```

```{r mc-hist, fig.height=3, out.width="88%", fig.align="center"}
mc_med <- median(mc_graded$mc_score)
ggplot(mc_graded, aes(x = mc_score)) +
  geom_histogram(binwidth = 1, fill = "#2c7fb8", color = "white", boundary = 0) +
  geom_vline(xintercept = mc_med, linetype = "dashed") +
  scale_x_continuous(breaks = seq(0, MC_MAX, 3)) +
  labs(x = paste0("MC score (out of ", MC_MAX, ")"), y = "Count",
       caption = paste0("Dashed line = median (", mc_med, ")")) +
  theme_minimal(base_size = 11)
```

---

## FR Scores (out of `r FR_MAX`)

Based on **`r n_fr`** students with FR graded.
*Note: non-random sample — graded students likely skew higher than the full class.*

```{r fr-stats}
fr_med <- median(fr_graded$fr_score)
kable(make_stats(fr_graded$fr_score, label_max = FR_MAX, digits = 1),
      booktabs = TRUE, linesep = "", align = "lr",
      col.names = c("Statistic", "Value")) %>%
  kable_styling(latex_options = "hold_position", full_width = FALSE, position = "center")
```

```{r fr-hist, fig.height=3, out.width="88%", fig.align="center"}
ggplot(fr_graded, aes(x = fr_score)) +
  geom_histogram(binwidth = 2, fill = "#d95f02", color = "white", boundary = 0) +
  geom_vline(xintercept = fr_med, linetype = "dashed") +
  scale_x_continuous(breaks = seq(0, FR_MAX, 4)) +
  labs(x = paste0("FR score (out of ", FR_MAX, ")"), y = "Count",
       caption = paste0("Dashed line = median (", fr_med, ")")) +
  theme_minimal(base_size = 11)
```

---

## Combined Raw Score (out of `r EXAM_MAX`)

Based on **`r n_both`** students with both MC and FR graded.
Raw score $= 4 \times \text{MC} + \text{FR}$.

```{r raw-stats}
raw_med <- median(fe$raw_score)
kable(make_stats_dual(fe$raw_score, EXAM_MAX),
      booktabs = TRUE, linesep = "", align = "lrr") %>%
  kable_styling(latex_options = "hold_position", full_width = FALSE, position = "center")
```

```{r raw-hist, fig.height=3, out.width="88%", fig.align="center"}
raw_med_pct <- round(raw_med / EXAM_MAX * 100, 1)
ggplot(fe, aes(x = raw_pct)) +
  geom_histogram(binwidth = 5, fill = "#756bb1", color = "white", boundary = 0) +
  geom_vline(xintercept = raw_med_pct, linetype = "dashed") +
  labs(x = "Raw score (%)", y = "Count",
       caption = paste0("Dashed line = median (", raw_med_pct, "%)")) +
  theme_minimal(base_size = 11)
```

---

## Curve

Scores are curved using an affine transformation anchored at two points:
the observed median raw score maps to **`r TARGET_MEDIAN`** and the observed
maximum maps to **`r TARGET_MAX`**.

$$B = \frac{`r TARGET_MAX` - `r TARGET_MEDIAN`}{`r raw_max` - `r raw_median`} = `r round(B_param, 4)`$$

$$A = `r TARGET_MEDIAN` - `r round(B_param, 4)` \times `r raw_median` = `r round(A_param, 4)`$$

$$\text{Curved score} = `r round(A_param, 2)` + `r round(B_param, 4)` \times \text{raw score}$$

**Cap rule:** curved score $\leq \min\!\bigl(\text{raw\%} + 0.25 \times \text{raw\%},\; 100\bigr)$.

```{r curve-plot, fig.height=3, out.width="88%", fig.align="center"}
curve_df <- tibble(raw = seq(0, EXAM_MAX)) %>%
  mutate(
    curved = A_param + B_param * raw,
    cap    = pmin(raw / EXAM_MAX * 100 * 1.25, 100),
    curved = pmin(curved, cap),
    curved = pmin(curved, 100),
    curved = pmax(curved, 0)
  )
ggplot(curve_df, aes(x = raw / EXAM_MAX * 100, y = curved)) +
  geom_line(color = "#e6550d", linewidth = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", color = "grey50") +
  geom_point(data = fe,
             aes(x = raw_pct, y = curved_score),
             alpha = 0.4, size = 1.2, color = "#2c7fb8") +
  labs(x = "Raw score (%)", y = "Curved score",
       caption = "Red = curve; blue dots = student scores; dotted = no adjustment") +
  theme_minimal(base_size = 11)
```

---

## Curved Scores (out of 100)

```{r curved-stats}
kable(make_stats(fe$curved_score, label_max = 100),
      booktabs = TRUE, linesep = "", align = "lr",
      col.names = c("Statistic", "Value")) %>%
  kable_styling(latex_options = "hold_position", full_width = FALSE, position = "center")
```

```{r curved-hist, fig.height=3, out.width="88%", fig.align="center"}
cur_med <- median(fe$curved_score)
ggplot(fe, aes(x = curved_score)) +
  geom_histogram(binwidth = 5, fill = "#41ab5d", color = "white", boundary = 0) +
  geom_vline(xintercept = cur_med, linetype = "dashed") +
  labs(x = "Curved score", y = "Count",
       caption = paste0("Dashed line = median (", cur_med, ")")) +
  theme_minimal(base_size = 11)
```

---

## Students with No Score / Pending

```{r pending}
all_ids <- mc %>% select(sis_user_id, mc_status)
pending <- all_ids %>%
  left_join(fr %>% select(sis_user_id, fr_status), by = "sis_user_id") %>%
  filter(!(mc_status == "Graded" & fr_status == "Graded")) %>%
  arrange(mc_status, fr_status)

if (nrow(pending) == 0) {
  cat("All enrolled students have been fully graded.\n")
} else {
  kable(pending, booktabs = TRUE, linesep = "", align = "lcc",
        col.names = c("SIS ID", "MC status", "FR status")) %>%
    kable_styling(latex_options = c("hold_position", "scale_down"),
                  full_width = FALSE, position = "center")
}
```

---

## Flags and Action Items

- **`r nrow(pending)` student(s) not yet fully graded** — re-run pipeline once FR grading is complete.
- **`r if (PRELIMINARY) "Curve is PRELIMINARY."` Recalibrate `TARGET_MEDIAN` in `03_combine_and_curve.R` once all FR scores are in.**
